Analyzing Trends in U.S. Residential Real Estate Prices

### 2024-04-15 outputs/paper/paper.qmd
paper updated

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-pos: 'h'
#| label: fig-trend_overall
#| fig-cap: Trend of Average House Price from 2000 to 2024
#| fig-width: 7
#| fig-height: 3.5

@@

@fig-trend_overall illustrates the average house price trend from 2000 to 2024. This graph highlights several inflection points in housing prices over the span of nearly a quarter of a century. A significant peak occurs in 2006, followed by a notable dip culminating in a trough in 2012. Subsequently, prices climb, reaching another prominent peak in 2022. The chart clearly marks these fluctuations, drawing attention to the cyclic nature of the housing market over time. When we ignore these brief highs or lows, the overall trend of housing prices is a clear upward trend.

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-pos: 'h'
#| label: fig-trend_by_type
#| fig-cap: Trend of Average House Price from 2000 to 2024 by House Type
#| fig-width: 7
#| fig-height: 3.5

@@

@fig-trend_by_type segments the average house price trends from 2000 to 2024 by house type, from one-bedroom to houses with five or more bedrooms. The trajectory for each house type is distinct, with the five-plus-bedroom category reaching the highest average prices, particularly exhibiting a sharp rise after 2012. The four-bedroom category shows substantial growth, albeit less sharply than the largest homes. One and two-bedroom homes reveal a more gradual increase, suggesting a different progression in price trends compared to larger properties. The disparities among the trends of different house types highlight the varied performance within the housing market. The overall price trend of different bedroom types is also showing a clear upward trend.

@@

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-pos: 'h'
#| label: fig-hm_overall
#| fig-cap: Average Price by State in the US for All House Types

@@

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-pos: 'h'
#| label: fig-hm_by_type
#| fig-cap: Average Price by State in the US for Different House Types
#| fig-subcap: ["one bedroom","two bedrooms","three bedrooms","four bedrooms","five plus bedrooms"]
#| layout-ncol: 2
#| layout-nrow: 3
#| fig-width: 7
#| fig-height: 3

@@

@@

$$
TBD
$$

@@

$$
TBD
$$

The Generalized Additive Model (GAM) represents a flexible approach to modeling complex, non-linear relationships between the dependent and independent variables. The setup for the GAM model in this study involves specifying smooth functions of the predictors (likely year and number of bedrooms), allowing for the accommodation of non-linear trends that cannot be captured by traditional linear models. Unlike polynomial regression, which explicitly models non-linearity through polynomial terms, GAM uses smoothing functions that are fitted to the data in a non-parametric manner. This approach enables the model to adapt to the shape of the underlying trend without pre-specifying the form of the relationship, thus providing a tailored fit to the peculiarities in the data.

$$
TBD
$$

@@

@@

@@

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-pos: 'h'
#| label: fig-GAM
#| tbl-cap: GAM Regression Fit Result

@@

>>>>

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-pos: 'h'
#| label: fig-trend
#| fig-subcap: ["Trend of Average House Price from 2000 to 2024", "Trend of Average House Price from 2000 to 2024 by House Type"]
#| layout-ncol: 1
#| layout-nrow: 2
#| fig-width: 7
#| fig-height: 4

@@

@@

@fig-trend-1 illustrates the average house price trend from 2000 to 2024. This graph highlights several inflection points in housing prices over the span of nearly a quarter of a century. A significant peak occurs in 2006, followed by a notable dip culminating in a trough in 2012. Subsequently, prices climb, reaching another prominent peak in 2022. The chart clearly marks these fluctuations, drawing attention to the cyclic nature of the housing market over time. When we ignore these brief highs or lows, the overall trend of housing prices has a clear upward trend.

@fig-trend-2 segments the average house price trends from 2000 to 2024 by house type, from one-bedroom to houses with five or more bedrooms. The trajectory for each house type is distinct, with the five-plus-bedroom category reaching the highest average prices, particularly exhibiting a sharp rise after 2012. The four-bedroom category shows substantial growth, albeit less sharply than the largest homes. One and two-bedroom homes reveal a more gradual increase, suggesting a different progression in price trends compared to larger properties. The disparities among the trends of different house types highlight the varied performance within the housing market. The overall price trend of different bedroom types is also showing a clear upward trend.

@@

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-pos: 'h'
#| label: fig-hm_overall
#| fig-cap: Average Price by State in the US for All House Types
#| fig-width: 7
#| fig-height: 2.5

@@

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-pos: 'h'
#| label: fig-hm_by_type
#| fig-cap: Average Price by State in the US for Different House Types
#| fig-subcap: ["one bedroom","two bedrooms","three bedrooms","four bedrooms","five plus bedrooms"]
#| layout-ncol: 2
#| layout-nrow: 3
#| fig-width: 6.5
#| fig-height: 3

@@

$$\text{Price} = \beta_0 + \beta_1 \times \text{Year} + \beta_2 \times \text{NumBedroom}$$

@@

$$\text{Price} = \beta_0 + \beta_1 \times \text{Year} + \beta_2 \times \text{Year}^2 + \beta_3 \times \text{NumBedroom} + \beta_4 \times \text{NumBedroom}^2$$

@@

$$\text{Intercept for GAM} = \beta_0$$

The Generalized Additive Model (GAM) represents a flexible approach to modeling complex, non-linear relationships between the dependent and independent variables. This Model does not really have a fixed equation with coefficients due to its use of smoothing functions. The setup for the GAM model in this study involves specifying the smooth functions of the predictors, year and number of bedrooms, allowing for the accommodation of non-linear trends that cannot be captured by traditional linear models. Unlike polynomial regression, which explicitly models non-linearity through polynomial terms, GAM uses smoothing functions that are fitted to the data in a non-parametric manner. This approach enables the model to adapt to the shape of the underlying trend without pre-specifying the form of the relationship, thus providing a tailored fit to the peculiarities in the data.

@@

After some understanding of model choices, @tbl-linear_model_result and @tbl-nonlinear_model_result displays the regression results, which will be explain with detials in later sections.

$$\text{Price} = -13,969,628.019 + 6,958.769 \times \text{Year} + 59,897.327 \times \text{NumBedroom}$$

$$\text{Price} = 212,976.031 + 3,866,485.706 \times \text{Year} + 1,601,816.793 \times \text{Year}^2$$
$$+ 6,591,537.312 \times \text{NumBedroom} + 1,365,656.076 \times \text{NumBedroom}^2$$

@@

$$\text{Intercept for GAM} = 212,976.031$$

@@

Furthermore, two charts in @fig-GAM_fit each showing the smoothed relationship between the predictor and the response variable (house price), with a solid line representing the estimated function and dashed lines indicating the confidence interval around the estimate.

The plot on the left shows the relationship between 'Year' and house prices. There’s a clear non-linear trend. The prices appear to have a cyclical pattern with a significant peak around 2006, followed by a sharp decline, which likely corresponds to the global financial crisis of 2008. After this trough, there’s a recovery with prices trending upwards again, but with a suggestion of leveling off towards the end of the period. This could indicate a stabilization in the market or potentially the beginning of another downturn.

The plot on the right shows the relationship between 'NumBedroom' and house prices. This relationship is more straightforward: as the number of bedrooms increases, so does the house price. The upward trend appears to be exponential rather than linear, suggesting that additional bedrooms have an increasingly positive effect on house prices. The confidence interval widens with the number of bedrooms, indicating more variability in the data for houses with more bedrooms, which could be due to fewer observations in the higher bedroom categories or greater variance in prices for larger homes.

Both plots reveal the flexibility of GAMs in modeling complex relationships, allowing for the identification of trends that simpler linear models could miss. The confidence intervals suggest the model is fairly confident about its estimates, especially for the 'NumBedroom' variable. However, the wider intervals at the extremes of the data range would call for cautious interpretation in those areas.

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-pos: 'h'
#| label: fig-GAM_fit
#| tbl-cap: GAM Regression Fit Result

@@

However, the visual representation of data trends provided by @fig-GAM_fit is not just intuitively appealing but also offers practical insights that might be obscured in tabular regression outputs. For example, the cyclical pattern observed in the 'Year' graph aligns with known economic events that affected housing prices, demonstrating GAM's capacity to capture externalities and temporal dynamics beyond the reach of conventional linear models.
